\documentclass[a4paper,11pt]{article}

% Padding Requirements
\usepackage[a4paper, margin=1in]{geometry}
% Used for images
\usepackage{graphicx}
\usepackage{float}
\graphicspath{{/Users/sloth_mini/Documents/4.1/dissertation/images/}}
% Used for bibliography
\usepackage[style=ieee, sorting=none, backend=biber]{biblatex}
\renewcommand{\bibfont}{\small}

% References file
\addbibresource{references.bib}


% ---- UNIVERSITY NAME ---- %
\newcommand{\uni}{Concordia University}
\newcommand{\city}{Montreal}
\newcommand{\program}{The Master of Applied Computer Science}
% ----- SUPERVISOR ----- %
\newcommand{\supervisor}{\textbf{Professor Neamat El Gayar}}
 
% ---- DISSERTATION TITLE ---- %
\newcommand{\project}{Multi-Element Association Rule Derivation from Minimum Spanning Trees}

% ----- HERIOT WATT UNVIERSITY (DUBAI?) ---- %
\newcommand{\hw}{Heriot-Watt University}

\begin{document}
\title{Mitigating Circumstances}

\author{Sahil Pattni\\(sp72@hw.ac.uk)}
\date{}
\maketitle

% I would like to reassure you that I am not that person anymore. I do this not with empty words, but my academic performance these past four years.
\section*{Heriot-Watt University [P grades]}
Towards the end of the second semester of my third academic year, the world was hit with the Covid-19 pandemic, and as universities, schools and businesses were closed under a government-mandated lockdown, my university did not have enough time to transition to an online examination system. Therefore, for courses whose grades were primarily determined through exams, a passing grade ($P$) was awarded since the exams could not be administered. For courses that were marked through courseworks, the regular grading system was used as all courseworks had been submitted by the time the lockdown was put into effect.

\section*{Purdue University}
In the early months of 2016,  I applied to Purdue University for a BSc. in Computer Science.  Unfortunately, I applied a little later than most people, and upon receiving my acceptance letter, I was informed that the computer science department was at their maximum capacity. Therefore, I was to be admitted to the Exploratory Studies program\footnote{Equivalent to an Undecided degree} with the option to transfer into Computer Science if my GPA was kept in good standing. 
Up until that point, I never had to work too hard to maintain a decent grade, and I ended up carrying that mindset as I entered Purdue University. As happy as I was to delve deeper into the topic of computer science, my torpid outlook towards my academics meant that I quickly fell behind on my classes. I realized that this lethargic mentality was a slippery slope of steadily declining mental health that led to a very low point in my life where I had lost the motivation to work or be productive; something that is represented quite accurately in my transcripts. However, the one thing I had never lost an interest in was computer science, and even at the height of my depression, I managed to perform well in that one course. However, one course could not save me from the many, and after my second semester I was informed that I could no longer continue to study at Purdue.
Dropping out of university has been the most personally challenging experience of my life, certainly the most impactful. The weight of this impact has also caused me to reflect heavily on my mistakes and has strengthened my resolve to succeed. With this new outlook, I've joined Heriot-Watt University in Dubai - pursuing my BSc. in Computer Science - and in the face of the same academic hurdles I faced four years ago, I've maintained a GPA of 3.77. I've worked very hard to get to where I am today, I know the challenges moving forward will be anything but easier, and I know I can carry my new-found work ethic over to \uni\ and succeed.

\end{document}.